---
title: "Data Preparation"
author: "Zachary Levonian"
date: "11/02/2018"
output: pdf_document
---

```{r}
library(alr4)
library(mice)  # for multiple imputation
```

```{r}
train <- read.csv("../../data/raw/train.csv", stringsAsFactors=FALSE)
test <- read.csv("../../data/raw/test.csv", stringsAsFactors=FALSE)
```

```{r}
test$Survived = 2
df <- rbind(train, test)
```


# Build factors from data

```{r}
df$fSex = factor(df$Sex)
df$fEmbarked = factor(df$Embarked)
```

Now, we save all the columns to be used as potential features to a file.

```{r}
toWrite <- df[,c("PassengerId", "fSex", "fEmbarked", "Pclass", "Age", "SibSp", "Parch", "Fare", "Survived")]
write.table(toWrite, file="../../data/derived/factorized_data.csv",
            row.names=FALSE, col.names=TRUE, 
            sep=",", quote=FALSE)
```


# Data Exploration

```{r}
pairs(df[c("fSex", "fEmbarked", "Pclass", "Age", "SibSp", "Parch", "Fare", "Survived")])
```


# Train a Model

The code below demonstrates:
 - Reading in the features
 - Splitting the data into the training and test data
 - Training a regression model
 - Predicting the test set based on the trained model
 - Saving the predictions in the Kaggle format for submission

```{r}
df <- read.csv("../../data/derived/factorized_data.csv", stringsAsFactors=TRUE)

```


```{r}
train <- df[df$Survived != 2, ]
test <- df[df$Survived == 2, ]
```

```{r}
md <- glm(Survived ~ fSex + fEmbarked + Pclass + Age + SibSp + Parch + Fare, family="binomial", data=train)
summary(md)
```


```{r}
# predict on the test set
y_pred <- predict(md, test, type="response")
hist(y_pred)
```

```{r}
# save the predictions to a file that can be submitted to Kaggle
test$Survived = as.numeric(y_pred > 0.5)
# while there shouldn't be NAs in the test output, 
# here we apply a nasty hack to ensure any NAs are numeric in the output
test[is.na(test)] <- 0
toWrite <- test[,c("PassengerId", "Survived")]
write.table(toWrite, file="../../data/derived/testModelPredictions.csv", 
            row.names=FALSE, col.names=TRUE, 
            sep=",", quote=FALSE)
```

